---
title: "US N2O data comparison"
author: "Dr. Chih-Yu Hung"
date: "2024-04-18"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
```

## Check if the N2O emission in the US has similar non-linear curve with Canada

```{r US data wrangling}
US_data <- read.csv("Input/US_N2OEF.csv")
CA_data <- read.csv("Input/Rochette_2018_PvsEF.csv") %>% #The data from Rochette 2018 Fig 3
           mutate(texture = Drain2)
  #US data wrangling
US_data <- US_data %>%
  filter (Crop.type == "Annual") %>%
  filter (Application.amount > 0) %>%
  filter (EF > 0) %>%
  filter (Precipitation > 0)

#Convert soil texture to three texture system
US_data <- US_data %>%
  mutate(texture =
  case_when(tolower(Soil.type) %in% c("clay", "silty clay", "silty clay loam", "clay loam", "sandy clay") ~ "Fine",
            tolower(Soil.type) %in% c("sandy loam","sandy clay loam","loam soil", "loam", "silt loam", "silt") ~ "Medium",
            TRUE ~ "Coarse")
  )


US_table <- US_data %>%
  group_by(Reference, Pub.year) %>%
  mutate(Reference = paste0(sub(" .*", "", Reference), " et al.")) %>%
  summarise(
    State = str_c(unique(State), collapse = ", "),
    Region = str_c(unique(Region), collapse = ", "),
    texture = str_c(unique(texture), collapse = ", "),
    crop = str_c(unique(Crop), collpase =", ")
  ) %>%
  ungroup()

#write.csv(US_table,"Output/Study summary.csv",row.names = F)

```

## Prepare the function used for analysis

The goal is to see if we can use Canada’s non linear model in the US 1. see if the Canada’s nonlinear model (y = e(0.00558*x-7.701)) fit US data I’ll use Mean sqaure error (MSE) and Symmetric Mean Absolute Percentage Error (SMAPE) 2. Plot a figure to show the dots from CA (hollow) and the dots from US (Solid), soil texture will be Fine (circle), Med (triangle), Coarse (square) 3. Make a table to compare using CA model coefficients and coefficients developed based on US data and their MSE and SMAPE 4. Plot a figure to show the relationship for predict EF values and published values

```{r integrated model, echo=FALSE}
integrated_model <- function(data, a_start, b_start) {
# Fit the model using non-linear least squares (NLS)
fit <- nls(EF ~ exp(a * Precipitation + b),data = data,
start = list(a = a_start, b = b_start))
# Extract fitted coefficients
a_fit <- coef(fit)[1]
b_fit <- coef(fit)[2]
# Calculate R-squared
residuals <- residuals(fit)
SS_RES <- sum(residuals^2)
SS_RES <- sum(residuals^2)
SS_TOT <- sum((data$EF - mean(data$EF))^2)
r_squared <- 1 - (SS_RES / SS_TOT)
# Generate predictions based on the fitted model
data$predictions <- exp(a_fit * data$Precipitation + b_fit)
data$predictions_CA <-exp(0.00558 * data$Precipitation + (-7.701))
# Calculate RMSE
rmse <- sqrt(mean((data$predictions - data$EF)^2))
rmse_CA <- sqrt(mean((data$predictions_CA - data$EF)^2))
# Calculate sMAPE
smape <- mean(abs(data$EF - data$predictions) / ((abs(data$EF) + abs(data$predictions)))) * 100
smape_CA <- mean(abs(data$EF - data$predictions_CA) / ((abs(data$EF) + abs(data$predictions_CA)))) * 100
# Create a result table
results_table <- tibble(
Metric = c("R-squared", "RMSE","RMSE to CA", "sMAPE", "sMAPE to CA"),
Value = c(round(r_squared, 4), round(rmse, 4),round (rmse_CA,4), round(smape, 2), round(smape_CA,2))
)
# Print the results
print(results_table)
# Plot with shapes based on "texture" and "Region" column
texture_shapes <- case_when(
data$texture == "Fine" & data$Region == "North" ~ 1,  # Circle for Fine and North
data$texture == "Fine" & data$Region == "South" ~ 16,  # Filled circle for Fine and South
data$texture == "Medium" & data$Region == "North" ~ 2,  # Triangle for Medium and North
data$texture == "Medium" & data$Region == "South" ~ 17,  # Filled triangle for Medium and South
data$texture == "Coarse" & data$Region == "North" ~ 0,  # Square for Coarse and North
data$texture == "Coarse" & data$Region == "South" ~ 15,  # Filled square for Coarse and South
TRUE ~ 0  # Default shape for other combinations
)
plot(data$Precipitation, data$EF,
pch = texture_shapes,
cex  = 2,
col = "blue",
xlab = "Precipitation",
ylab = "EF",
main = "Fitted Exponential Model with Texture",
xlim = c(0, max(data$Precipitation)),
ylim = c(0, max(data$EF)))
# Add fitted curve
curve(exp(a_fit * x + b_fit), from = min(data$Precipitation), to = max(data$Precipitation),
add = TRUE, col = "red", lwd = 2)
curve(exp(0.00558 * x + (-7.701)), from = min(data$Precipitation), to = max(data$Precipitation),
add = TRUE, col = "black", lwd = 2, lty = "dotted")
# Add legend for texture types
legend("topleft", legend = c("Fine", "Medium", "Coarse"),
col = "blue", pch = c(1, 2, 0), title = "Texture")
#Observation vs prediction plot
plot(data$EF, data$predictions,
pch = texture_shapes,
cex  = 2,
col = "blue",
xlab = "Observation",
ylab = "Prediction",
main = "Observation vs. prediction",
xlim = c(0, max(data$EF)),
ylim = c(0, max(data$EF)))
# Add 1:1 curve
abline(a = 0, b = 1)
# Add legend for texture types
legend("topleft", legend = c("Fine", "Medium", "Coarse"),
col = "blue", pch = c(1, 2, 0), title = "Texture")


# Return the fit and table as output
return(list(fit = fit, results = results_table))




}

```

```{r CA data}

integrated_model(CA_data, a_start = 0.005, b_start = -7.7)

```


```{r US data}
integrated_model(US_data, a_start = 0.005, b_start = -7.7)

#Divided to south and north
US_South <- US_data %>%
  filter(Region == "South")

US_North <- US_data %>%
  filter(Region == "North")


integrated_model(US_North, a_start = 0.005, b_start = -7.7)
integrated_model(US_South, a_start = 0.005, b_start = -7.7)
```


```{r US data prediction vs observation}



```